---
title: "Untitled"
output: word_document
---

```{r}
setwd("C:/I")
library(RSQLite)
imbd = dbConnect(SQLite(), "lean_imdbpy.db")

# look tables in the data
dbListTables(imbd)
```


**1.How many actors are there in the database? How many movies?**

```{r Q1}
# rows in name table
dbGetQuery(imbd, "SELECT COUNT(*) AS NumberofActors From name")

# rows in title table, limit kind = "movie"
dbGetQuery(imbd, "SELECT COUNT(*) AS Numberofmovies
                  From title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                  WHERE kind = 'movie'") 
 
```

**2.What time period does the database cover?**

```{r}
# look at title table (won't show the output in the homwwork)
#dbGetQuery(imbd, "SELECT * From title LIMIT 5")
```

```{r}
# get max year
dbGetQuery(imbd, "SELECT MAX(production_year) From title")

# get MIN year
dbGetQuery(imbd, "SELECT MIN(production_year) From title")

```

**3.What proportion of the actors are female? male?**

```{r}
# counts by gender and devide by the total count in name
dbGetQuery(imbd, "SELECT gender, count(*)*100 / (SELECT COUNT(*) 
                                                 FROM name) 
                  AS proportion
                  From name
                  GROUP BY gender")
```

**4. What proportion of the entries in the movies table are actual movies and what proportion are television series, etc.?**
This one is similar like question 3, so I use the same idea to do it, just replace gender by kind and use the combine of title and kind_type table

```{r}
dbGetQuery(imbd, "SELECT kind, COUNT(*)*100 / (SELECT COUNT(*) 
                                               FROM title) 
                  AS proportion
                  From title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                  GROUP BY kind")

```

**5.How many genres are there? What are their names/descriptions?**

```{r}
# look at info_type table (won't show it in the homework)
#dbGetQuery(imbd, "SELECT *  From info_type LIMIT 5")
```

```{r}
# distinct genres type, get genres' names
genre = dbGetQuery(imbd, "SELECT DISTINCT info  
                          From movie_info 
                          WHERE info_type_id = 3 ")
genre[,1]
```

**6.List the 10 most common genres of movies, showing the number of movies in each of these genres.**

```{r}
dbGetQuery(imbd, "SELECT movie_info.info, count(*) AS Frequncy 
                  From title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN movie_info 
                            ON title.id = movie_info.movie_id
                  WHERE info_type_id = 3 
                  AND kind_type.kind = 'movie'
                  GROUP BY movie_info.info
                  ORDER BY Frequncy DESC 
                  LIMIT 10")


```



**7.Find all movies with the keyword 'space'. How many are there? What are the years these were released? and who were the top 5 actors in each of these movies?**

```{r}
# count the number of movies where keyword = space
dbGetQuery(imbd, "SELECT count(*) AS MovieswithSpace 
                  FROM title 
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN movie_keyword 
                             ON movie_keyword .movie_id = title.id
                       JOIN keyword
                             ON keyword.id = movie_keyword.keyword_id
                  WHERE keyword.keyword = 'space' 
                  AND kind = 'movie'")
```

For the years of these movie, I use the same way as above, I only change the return value to be production_year.

```{r}
year = dbGetQuery(imbd, "SELECT production_year 
                         FROM title 
                              JOIN kind_type   
                                   ON title.kind_id = kind_type.id
                              JOIN movie_keyword 
                                   ON movie_keyword .movie_id = title.id
                              JOIN keyword
                                   ON keyword.id = movie_keyword.keyword_id
                          WHERE keyword.keyword = 'space'
                          AND kind = 'movie' ")
year[,1]
```

```{r}

# creat keyword and movie_id temporary table
dbGetQuery(imbd, "CREATE TEMPORARY TABLE space_movie AS 
                  SELECT  title.id, title.title
                  FROM title 
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN movie_keyword 
                             ON movie_keyword.movie_id = title.id
                       JOIN keyword
                             ON keyword.id = movie_keyword.keyword_id
                  WHERE keyword.keyword = 'space'
                  AND kind = 'movie'")

# get the top five actors name, billing position in 1 to 5 for each movies. 
dbGetQuery(imbd, "SELECT  name.name, cast_info.nr_order, space_movie1.title
                  FROM space_movie1
                       JOIN title
                             ON title
                       JOIN cast_info
                             ON space_movie1.id = cast_info.movie_id
                       JOIN name
                             ON cast_info.person_id = name.id
                  WHERE cast_info.nr_order IN (1,2,3,4,5) 
                  LIMIT 5 ")



```

 
** 8.Has the number of movies in each genre changed over time? Plot the overall number of movies in each year over time, and for each genre.**
 

```{r}
num_movie = dbGetQuery(imbd, "SELECT production_year, info
                  From movie_info 
                       JOIN title
                            ON title.id = movie_info.movie_id
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                  WHERE info_type_id = 3 
                  AND kind = 'movie'
                  AND production_year IS NOT NULL
                  ORDER BY production_year")

# use ggplot to draw the number of movies in each year
library(ggplot2)
ggplot(num_movie, aes(x=production_year , fill=info )) + geom_bar() + ggtitle("Genre by year")

```

**#9 Who are the actors that have been in the most movies? List the top 20.**

```{r}
dbGetQuery(imbd, "SELECT  name, COUNT(*) AS appear 
                  FROM title 
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN cast_info 
                             ON cast_info.movie_id = title.id
                       JOIN name
                             ON name.id = cast_info.person_id
                  WHERE kind = 'movie'
                  GROUP BY name
                  ORDER BY appear DESC
                  LIMIT 20")
```

** 10 Who are the actors that have had the most number of movies with "top billing", i.e., billed as 1, 2 or 3? For each actor, also show the years these movies spanned?**

```{r}
# why need cast_info.movie_id
actor10 = sapply(1:3, function(x) { 
      y = paste0("SELECT  name, nr_order, cast_info.movie_id, COUNT(*) AS appear 
                  FROM title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN cast_info
                             ON title.id = cast_info.movie_id
                       JOIN name
                             ON cast_info.person_id = name.id
                  WHERE cast_info.nr_order = ",
                  x,
                  "
                  AND kind = 'movie'
                  GROUP BY name 
                  ORDER BY appear DESC
                  LIMIT 5")
      dbGetQuery(imbd,y)
                               })
```


```{r}
year = sapply(1:3, function(x) {
            y = paste0("SELECT name, nr_order, cast_info.movie_id, MAX(production_year), COUNT(*) AS appear 
                  FROM title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN cast_info
                             ON title.id = cast_info.movie_id
                       JOIN name
                             ON cast_info.person_id = name.id
                  WHERE cast_info.nr_order = ",
                  x,
                  "
                  AND kind = 'movie'
                  GROUP BY name 
                  ORDER BY appear DESC
                  LIMIT 5")
            dbGetQuery(imbd,y)
                             })
```

```{r}
dbGetQuery(imbd,"SELECT name, nr_order, cast_info.movie_id, MAX(production_year), COUNT(*) AS appear 
                  FROM title
                       JOIN cast_info
                             ON title.id = cast_info.movie_id
                       JOIN name
                             ON cast_info.person_id = name.id
                  WHERE cast_info.nr_order = 1
                  GROUP BY name 
                  ORDER BY appear DESC
                  LIMIT 5")
```

**11 Who are the 10 actors that performed in the most movies within any given year? What are their names, the year they starred in these movies and the names of the movies?**


```{r}
dbGetQuery(imbd, "CREATE TABLE question11 AS 
                  SELECT title.production_year, name.name, name.id, COUNT(*) AS appear 
                  FROM title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN cast_info
                             ON title.id = cast_info.movie_id
                       JOIN name
                             ON cast_info.person_id = name.id
                  WHERE kind = 'movie'
                  GROUP BY title.production_year, name.name
                  ORDER BY appear DESC
                  LIMIT 10")
```


```{r}
dbGetQuery(imbd, "SELECT MIN(title.production_year), name.name 
                  FROM title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN cast_info
                             ON title.id = cast_info.movie_id
                       JOIN question11
                             ON cast_info.person_id = name.id
                  WHERE kind = 'movie'
                  GROUP BY name.name")
```

```{r}
dbGetQuery(imbd, "SELECT title.title, name.name 
                  FROM title
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN cast_info
                             ON title.id = cast_info.movie_id
                       JOIN question11
                             ON cast_info.person_id = name.id
                  WHERE kind = 'movie'
                  GROUP BY name.name")
```


**12 Who are the 10 actors that have the most aliases**

```{r}
dbGetQuery(imbd, "SELECT name.name, count(*) AS appear
                  FROM aka_name
                       JOIN name
                            ON aka_name.person_id = name.id
                  GROUP BY aka_name.person_id
                  ORDER BY appear DESC
                  LIMIT 10")
```

**13Networks: Pick a (lead) actor who has been in at least 20 movies. Find all of the other actors that have appeared in a movie with that person. For each of these, find all the people they have appeared in a move with it. Use this to create a network/graph of who has appeared with who. Use the igraph or statnet packages to display this network. **

```{r}
# join the title, cast_info, and name table, find the actors name and id who have been in exact 21 movies.
 dbGetQuery(imbd, "SELECT name.id, name.name, COUNT(*) AS appear 
                  FROM title 
                       JOIN kind_type   
                            ON title.kind_id = kind_type.id
                       JOIN cast_info
                            ON title.id = cast_info.movie_id
                       JOIN name
                            ON cast_info.person_id = name.id
                  WHERE kind = 'movie'
                  GROUP BY name.name
                  HAVING appear = 21                
                  ORDER BY appear DESC
                  LIMIT 5 ")


       id                name appear
1 4199871 A Ljungberg, Markus     21
2     913        Aalam, Steve     21
3 2261940      Aaltonen, Sari     21
4 4635240     Aaron, David L.     21
5    1282         Aaron, Mark     21




 dbGetQuery(imbd, " CREATE TABLE movie_id AS 
                    SELECT title.id 
                    FROM title 
                         JOIN cast_info
                              ON title.id = cast_info.movie_id
                         JOIN name
                              ON cast_info.person_id = name.id
                    WHERE name.id = 913")


 dbGetQuery(imbd, "CREATE TABLE first_name AS 
                    SELECT DISTINCT title.id, name.name, name.id
                    FROM title 
                         JOIN movie_id
                              ON title.id = movie_id.id
                         JOIN cast_info
                              ON movie_id.id = cast_info.movie_id
                         JOIN name
                              ON cast_info.person_id = name.id ")

 

 dbGetQuery(imbd, "CREATE TABLE second_movie AS 
                                SELECT DISTINCT title.id 
                                FROM title 
                                      JOIN cast_info
                                           ON title.id = cast_info.movie_id
                                      JOIN name
                                           ON cast_info.person_id = name.id
                                      JOIN first_name
                                           ON first_name.id = name.id")




 dbGetQuery(imbd, "CREATE TABLE second_name AS
                                SELECT DISTINCT title.id, name.name, name.id, cast_info.nr_order
                                FROM second_movie
                                      JOIN cast_info
                                            ON second_movie.id = cast_info.movie_id
                                      JOIN name
                                            ON cast_info.person_id = name.id ")

second_movie =  dbGetQuery(imbd, "SELECT * FROM second_movie")
first_name =  dbGetQuery(imbd, "SELECT * FROM first_name")
second_name =  dbGetQuery(imbd, "SELECT * FROM second_name")

names(first_name) = c("name", "name.id", "title.id")
names(second_name) = c("title.id", "name", "name.id", "nr_order")

top_bill = second_name[second_name[,4] == 1, ]
top_bill = na.omit(top_bill)
top_bill$to = NA


data = lapply(1:nrow(first_name), function(x) top_bill[top_bill[,1] == first_name[x,3],5] = first_name[x, 1])

for(i in 1:nrow(first_name))
{
    
    top_bill[top_bill[,1] == first_name[i,3],5] = first_name[i,1]
}

```
